Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2015-03-08 sas2015/content/background.tex
removing some old comments

A path in the control-flow supergraph is \textit{valid} if it follows the usual stack-based calling 
discipline: every end-to-return edge on the path returns
to the site of the most recent call that has not yet been matched by a return. 
%\ftodo{Can we explain/define the notion of the ``most recent unmatched call'' more precisely?}
The set of all valid paths from the program entry point to
a node $n$ is denoted $\ivp(n)$.
%\otodo{Should we give a formal definition?}

%Let each call node in a program be labeled with a distinct opening parenthesis and the corresponding return node with the matching closing parenthesis. For a given path $p$, let $s$ be the string that is obtained by concatenating the labels of the nodes in $p$.
%Then $p$ is \textit{valid} if $s$ belongs to the language of substrings of balanced parentheses. The set of all interprocedurally valid paths from the start node to a node $n$ is denoted as $\ivp(n)$. The set $\ivp(n)$ is a conservative approximation of all concrete execution paths from the start node to $n$.

%A \textit{context-sensitive} data-flow analysis is an analysis that considers only interprocedurally valid paths.

\commentout{
\begin{mdelete}
\begin{example}
  In the supergraph in Figure~\ref{fig:examplesupergraph}, let us assign $\{,\,\}$ parentheses to $\textsf{call}_\texttt{A.f}$ and $\textsf{return}_\texttt{A.f}$, and $\langle,\,\rangle$ parentheses to $\textsf{call}_\texttt{f}$ and $\textsf{return}_\texttt{f}$. 
  Then the string corresponding to the path 
  \[
    p_1=\left[\highlight{\textsf{call}_\texttt{A.f}}{greyblue},\,
          \highlight{\textsf{start}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\texttt{if (s == null)}}{lightsalmonpink},\,
          \highlight{\texttt{return s}}{lightsalmonpink},\,
          \highlight{\textsf{end}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\textsf{return}_\texttt{A.f}}{greyblue}
    \right]
  \]
  is $\{\}$, which indicates that $p_1$ is valid. Every prefix of $p_1$ is also a valid path.

However, the graph also contains an interprocedurally invalid path
  \[
    p_2=\left[\highlight{\textsf{call}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\textsf{start}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\texttt{if (s == null)}}{lightsalmonpink},\,
          \highlight{\texttt{return s}}{lightsalmonpink},\,
          \highlight{\textsf{end}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\textsf{return}_\texttt{A.f}}{greyblue}
    \right]
  \]
  with corresponding string $\langle\}$.
\end{example}
\end{mdelete}
}

%\otodo{The definitions that Marianna had were for \emph{complete}
%lattices/semilattices. All of the lattices in the context of IFDS/IDE
%are/must be complete. I'll just say that here.}
%\ftodo{Would it make sense to first the define the notion of a lattice without referring to its completeness?
       %This could be followed by defining what it means for a (semi)lattice to be complete, and then we
       %could say that all lattices considered in this paper are complete, etc.}
%\ftodo{Can we define the symbols $\sqcup$ and $\sqcap$ for join and meet here?}
A \textit{lattice}%
\footnote{The definitions that we give here are of \textit{complete} lattices and semilattices.
Since all of the (semi)lattices discussed in this paper are required to be complete, we omit the
\textit{complete} qualifier.}
is a partially ordered set $(S,\sqsubseteq)$ in which every subset has a least upper bound, called \textit{join} or $\sqcup$, and a greatest lower bound, called \textit{meet} or $\sqcap$.
A \textit{meet semilattice} is a partially ordered set in which every subset only has a greatest lower bound.
The symbols $\bot$ and $\top$ are used to denote the greatest lower bound of $S$ and of the empty set, respectively.

%A \textit{meet semilattice} $L=(S,\,\sqcap)$ is defined by a set $S$ and a meet operation $\sqcap$ that is  associative, commutative, and idempotent.
  %The meet operation induces a partial order $(S,\,\sqsubseteq)$ where every subset contains  a greatest lower bound: For all $x,\,y\in S$, $x\sqsubseteq y$ if $x\sqcap y=x$. The greatest lower bound, or top element, of the semilattice is denoted as $\top$.
  %If $k$ is the length of the longest chains of elements in the semilattice, then the \textit{height} of the semilattice is~$k-1$.

In this paper, we denote a map $m$ as a set of pairs of keys and values, in which each key appears at most once.
For any map $m$,  $m(k)$ is the value paired with the key $k$ in $m$. We denote by $m[x\to y]$ a map that maps $x$ to $y$, and every other key $k$ to $m(k)$.
%To avoid excessive parentheses, we write $\left(m[x_1\to y_1]\right)[x_2\to y_2]$ as $m[x_1\to y_1][x_2\to y_2]$.

%We will denote the identity function $\lambda x\,.x$ by $\textsf{id}$.
%We will use a typed version of this function in various contexts, where the type of $x$ will vary with
%the context.

%OL: we can just as easily write the identity explicitly as \x.x throughout.

\subsection{IFDS}\label{sec:bgifds}
The IFDS framework~\cite{reps1995precise} is a precise and efficient algorithm for data-flow analysis
%. IFDS was developed in 1995 by T.\,Reps, S.\,Horwitz, and M.\,Sagiv at the University of Wisconsin and
that has been used to solve a variety of data-flow analysis problems~\cite{bodden2013spl,naeem2008typestate,DBLP:conf/birthday/KreikerRRSWY13,tripp2009taj}.
%\mtodo{ The four references take away space in the References section. Remove some of them?  }
The IFDS framework is an instance of the \textit{functional approach} to data-flow analysis~\cite{pnueli1981two}
because it constructs summaries of the effects of called procedures.
\commentout{The IFDS analysis is a version of the classic \textit{functional approach} to data-flow analysis proposed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}.}
\commentout{In other data-flow algorithms not based on the functional approach, the result of the analysis at the entry of a procedure ``merges'' the incoming data obtained from all callers of the procedure. As a consequence, there is one global data-flow result computed at the end of the procedure. Context-sensitivity, however, allows an analysis to compute the data-flow result for a given procedure as a \textit{function} of the data-flow value at the start of the procedure. 
The analysis result for a procedure varies depending on where the procedure was called from. This significantly improves the precision of a data-flow analysis, which is why context-sensitivity is an important advantage of IFDS over classic data-flow algorithms.}
The IFDS framework is applicable to \textit{interprocedural} data-flow problems whose domain is \textit{subsets} of a \textit{finite} set $D$,
and whose data-flow functions are \textit{distributive}.
A function $f$ is distributive if 
$f(x_1\sqcap x_2)=f(x_1)\sqcap f(x_2)$.

%The IFDS algorithm is applicable to problems which can be expressed with data-flow functions that satisfy certain restrictions. \textit{Interprocedural} flow functions specify how data flows from the invocation of a procedure to its start, and from the procedure's end back to its call site. \textit{Distributive} flow functions are those that distribute over the confluence operator. In the context of IFDS, the confluence operator is called meet, and it can be either union or intersection. The data-flow facts on which the analysis operates must be a \textit{finite} set $D$. Each flow function operates on a \textit{subset} of $D$ (for example, the set of variables in the program) which makes the domain of the flow functions the power set of $D$.

\commentout{The purpose of the IFDS framework is to solve a special subset of interprocedural, flow-sensitive, context-sensitive data-flow-analysis problems. The main idea of IFDS is to encode the data-flow analysis problem into a graph-reachability problem.}

@@

%\subsubsection{Data-Flow Problems Suitable for IFDS}\label{sec:suitableifds}
\commentout{In this section we describe the data-flow problems that can be solved by an IFDS analysis. We will start with an intuitive definition and later on formalize the notion of an IFDS-suitable problem.}

%Informally, an IFDS analysis can only solve decision problems. An IFDS analysis answers questions of the following kind: ``is property $X$ true at program point~$Y$?''. For example, a taint-analysis problem asks, for each variable $v$ in the program, ``is $v$ secret at a given program point?''. 
\commentout{An available-expressions problem asks, for each expression $e$, ``does $e$ have to be recomputed at a given program point?''.}

@@

\commentout{
is a function that maps supergraph edges to data-flow functions, and $M_F$ is extended to paths by composition\footnote{%
  Let $A$ be a set and $f:\,E^*\to(A\to A)$ a function from supergraph edges to functions on $A$.
  We say that $f$ is extended to  paths by composition to denote that for a path $q$ consisting of the edges $e_1,\,\dots,\,e_k$, $f(q)=f(e_k)\circ\ldots\circ f(e_1)\circ\id$.
}. The \textit{meet operator} $\sqcap$ is either union or intersection.
}

%Without loss of generality, we will take meet to denote union. It can be shown that any problem where meet is defined as intersection can be reformulated into an equivalent one where meet is defined as union~\cite{reps1995precise}.

@@

\commentout{
The representation relation allows us to compactly represent the composition and meet operations which are required for the IFDS algorithm.

For two representation relations $R_{f_1},\,R_{f_2}$, the composition and meet operations are defined as follows:
\begin{align}
  R_{f_1}\circ R_{f_2}&=\{(d_1,\,d_3)\ |\ \exists d_2:\ (d_1,\,d_2)\in R_{f_1},\,(d_2,\,d_3)\in R_{f_2}\}\,.
\intertext{and}
  R_{f_1}\sqcap R_{f_2}&=R_{f_1}\cup R_{f_2}\,.
\end{align}

The representation relation distributes over composition and meet:
\begin{align}
  R_{f_1}\circ R_{f_2}=R_{f_1\circ f_2}\,,\qquad
\text{and}\qquad
  R_{f_1}\sqcap R_{f_2}=R_{f_1\sqcap f_2}\,.
\end{align}
}

%On the exploded graph, the composition of two functions is represented by the paths that are formed when the exploded-graph edges are combined.
%\otodo{The example used to say $R_f \circ R_g$, but I believe it showed $R_g \circ R_f$. I fixed it, but please check that this is correct.}

@@

%\otodo{Discuss the micro-functions in the example exploded supergraph in Figure 3.}
\begin{example}
    In the exploded supergraph in Figure~\ref{fig:cc_edgefn_example},
    the micro-functions are shown as labels on the graph edges.
    Every edge without an explicit label has the identity as its
    micro-function. The micro-functions on the three edges from
    the node \verb+return secret()+ to the node $\texttt{end}_\texttt{A.foo}$
    together represent the environment transformer 
    $\lambda e.e[\psi \to \lambda m.\bot \sqcap \lambda m.m]$.

>>>>

A path in the control-flow supergraph is \textit{valid} if it follows the usual stack-based calling 
discipline: every end-to-return edge on the path returns
to the site of the most recent call that has not yet been matched by a return. 
%\ftodo{Can we explain/define the notion of the ``most recent unmatched call'' more precisely?}
The set of all valid paths from the program entry point to
a node $n$ is denoted $\ivp(n)$.

A \textit{lattice}%
\footnote{The definitions that we give here are of \textit{complete} lattices and semilattices.
Since all of the (semi)lattices discussed in this paper are required to be complete, we omit the
\textit{complete} qualifier.}
is a partially ordered set $(S,\sqsubseteq)$ in which every subset has a least upper bound, called \textit{join} or $\sqcup$, and a greatest lower bound, called \textit{meet} or $\sqcap$.
A \textit{meet semilattice} is a partially ordered set in which every subset only has a greatest lower bound.
The symbols $\bot$ and $\top$ are used to denote the greatest lower bound of $S$ and of the empty set, respectively.

In this paper, we denote a map $m$ as a set of pairs of keys and values, in which each key appears at most once.
For any map $m$,  $m(k)$ is the value paired with the key $k$ in $m$. We denote by $m[x\to y]$ a map that maps $x$ to $y$, and every other key $k$ to $m(k)$.

\subsection{IFDS}\label{sec:bgifds}
The IFDS framework~\cite{reps1995precise} is a precise and efficient algorithm for data-flow analysis
%. IFDS was developed in 1995 by T.\,Reps, S.\,Horwitz, and M.\,Sagiv at the University of Wisconsin and
that has been used to solve a variety of data-flow analysis problems~\cite{bodden2013spl,naeem2008typestate,DBLP:conf/birthday/KreikerRRSWY13,tripp2009taj}.
%\mtodo{ The four references take away space in the References section. Remove some of them?  }
The IFDS framework is an instance of the \textit{functional approach} to data-flow analysis~\cite{pnueli1981two}
because it constructs summaries of the effects of called procedures.
The IFDS framework is applicable to \textit{interprocedural} data-flow problems whose domain is \textit{subsets} of a \textit{finite} set $D$,
and whose data-flow functions are \textit{distributive}.
A function $f$ is distributive if 
$f(x_1\sqcap x_2)=f(x_1)\sqcap f(x_2)$.

@@

@@

@@

@@

\begin{example}
    In the exploded supergraph in Figure~\ref{fig:cc_edgefn_example},
    the micro-functions are shown as labels on the graph edges.
    Every edge without an explicit label has the identity as its
    micro-function. The micro-functions on the three edges from
    the node \verb+return secret()+ to the node $\texttt{end}_\texttt{A.foo}$
    together represent the environment transformer 
    $\lambda e.e[\psi \to \lambda m.\bot \sqcap \lambda m.m]$.
